---
title: 'Data Handling: Import, Cleaning and Visualisation'
author: "Dr. Aurélien Sallin"
output:
  html_document:
    highlight: tango
    theme: cerulean
    mathjax: "http://mathjax.rstudio.com/latest/MathJax.js?config=TeX-MML-AM_CHTML"
  pdf_document:
    pandoc_args:
    - --filter
    - ../../code/math.py
header-includes:
- \usepackage[T1]{fontenc}
- \usepackage{hyperref}
css: ../../style/notes_hsg.css
bibliography: ../../references/datahandling.bib
---

# Illustration Lecture 3

In this short illustration, we will see how to import an html content from the web. This will serve as an example to get an idea about a computing environment and the interaction between man and computer.

<br>

## Roadmap

  1. Access a website (over the Internet), use keyboard to enter data into a website. 

  2. R program accesses the data of the Google sheet (again over the Internet), downloads the data, and loads it into RAM.

  3. Data processing: produce output (in the form of statistics/plots), output on screen.
  
  
<br>

## Load necessary packages
```{r warning=FALSE}
library(httr)
library(pryr)
```

We want to load the website "https://www.economist.com/". We use the `GET()` function from the package `httr`.

```{r}
economist <- GET("https://www.economist.com/")
```

We see that we have three types of data: (1) the url written by hand in the form of a string, (2) the source code HTML being sent through the network, and (3) the HTML document loaded in the RAM.

We can ask the size of the object in the RAM using `object_size()` from package `pryr`.

```{r}
pryr::object_size(economist)
```


<br>

The html text data was sent in hexadecimal byte values to our computer. This is visible using the command `content` on our query, specifying that we want to see the raw data.
```{r}
economistRaw <- content(economist, as = "raw")
head(economistRaw, 15)
```
When translated to ASCII characters, it represents a string typically found at the beginning of HTML documents.

Hexadecimal  | Character 
-----|-----
3c | <
21 | !
44 | D
4f | O
43 | C
54 | T
59 | Y
50 | P
45 | E
20 | (a space character)
68 | h
74 | t
6d | m
6c | l

<br>

Now we can encode these hexadecimal values using the command `content(., as = "text")`.

```{r, collapse=TRUE}
economistText <- content(economist, as = "text")
head(economistText)
```
<br>

Finally, we can tell R to store the source code of the downloaded webpage into our local hard drive. 

```{r, collapse=TRUE}
writeLines(economistText, "economistText.html")
```