---
title: 'Data Handling: Import, Cleaning and Visualisation'
subtitle: "Exercise to lecture 4: csv and arrays"
author: "Dr. Aurélien Sallin"
output:
  html_document:
    highlight: tango
    theme: cerulean
    mathjax: "http://mathjax.rstudio.com/latest/MathJax.js?config=TeX-MML-AM_CHTML"
header-includes:
  -\usepackage[T1]{fontenc}
  -\usepackage{hyperref}
css: ../../style/notes_hsg.css
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
# knitr::opts_chunk$set(results = "hide")
```


# Working with a data frame

```{r data creation, include = FALSE, eval= TRUE}
library(readr)

# Data Creation (to be executed by the instructor)
set.seed(1234) # For reproducibility

firms <- c("FirmA", "FirmB", "FirmC", "FirmD", "FirmE")
years <- 2017:2022

# Generate sample data
data <- expand.grid(Firm = firms, Year = years)
data$Revenue <- sample(1000:5000, nrow(data), replace = TRUE)
data$Profit <- data$Revenue * runif(nrow(data), 0.1, 0.5)
data$Category <- factor(sample(c("Tech", "Health", "Finance"), nrow(data), replace = TRUE))
data$Revenue[10] <- "1933\xF6"
Encoding(data$Revenue) <- "latin1"

# Write data to .txt with ":" as separator
write.table(data, "financial_data.txt", row.names = FALSE, sep = ":",  fileEncoding = "latin1")

```


## Import data
Have a look at the file `financial_data.txt` using your favorite text editor. What do you notice? 

Import the table using the `read.csv()` function in your environment. Make sure you have the right path to access the .txt document. What does this parser do? Explore the data.frame. What is its structure? What are its dimensions?

Compute the summary statistics for each variable using the `summary()` command. What does this command give you? What do you notice? Make the necessary changes.


```{r eval = TRUE}
# Set correct path!

financial_data <- read.csv("financial_data.txt")
financial_data
```

```{r warning=FALSE}
# Import
iconv("\xF6", from = "ISO-8859-2", to = "UTF-8")
iconv("\xF6", from = "ISO-8859-1", to = "UTF-8")
iconv("\xF6", from = "ISO-8859-9", to = "UTF-8")

financial_data <- read.csv("financial_data.txt", fileEncoding = "ISO-8859-2")
financial_data
```

```{r}
financial_data <- read.csv("financial_data.txt", fileEncoding = "latin1", sep = ":")

LINE <- readLines("financial_data.txt")
head(read_lines_raw("financial_data.txt"))
```

```{r}
head(financial_data, 10)
summary(financial_data)

# Shift to numeric
financial_data[10, "Revenue"]
financial_data[10, "Revenue"] <- 1933

financial_data[, "Revenue"] <- as.numeric(financial_data[, "Revenue"])

# Check summary again
summary(financial_data)

```


```{r, echo = FALSE, eval=FALSE}
class(financial_data)
dim(financial_data)

str(financial_data)
names(financial_data)
dimnames(financial_data)
```

<br>

## Variable creation

Create a new variable "costs", which is the revenue - profit. [There are many ways to create a variable in a data frame. Here, use the `$` index.]

```{r, eval = TRUE, echo = FALSE}
financial_data$costs <- financial_data$Revenue - financial_data$Profit
```

<br>

## Factor variable

Which variable is (should be) a factor? Recode this variable as a factor. What are the levels? Should we have the variable `Firm` as a factor?

```{r, eval = FALSE, echo = TRUE}
financial_data$Category <- as.factor(financial_data$Category)

levels(financial_data$Category)
```

<br>

## Nests 
Split your data using the factor variable into three data frames that are contained in a list. Compute the mean profit for each data frame. Hint: use the function `split`.

```{r, eval = FALSE, echo = TRUE}
list_financial_data <- split(financial_data, financial_data$Category)

for (i in 1:length(list_financial_data)){
  print(mean(list_financial_data[[i]]$Profit))
}

# Or, using lapply
lapply(list_financial_data, function(x) mean(x$Profit))
```


<br>

## Advanced: map (not exam relevant)

Do the same as the exercise above using the `map` function. Install the packages `tidyr`, `dplyr`, and `purrr`.

```{r message=FALSE, warning=FALSE}
# Or (advanced!) with a nested tibble and map 
library(tidyr)
library(dplyr)
library(purrr)

tibble_financial_data <- financial_data |>
  group_by(Category) |>
  nest()

map(tibble_financial_data$data, ~mean(.$Profit))

```


<!-- <br> -->

<!-- <br> -->

<!-- # An exercise on arrays -->

<!-- ## Using the code below, create an array containing three cities and temperatures -->

<!-- Make sure you understand each line of code. This is exam relevant. Note: the function `rnorm(x)` randomly draws `x` values from a normal distribution with mean 0 and sd 1. We use it here to add some "random" variation in our simulated temperatures.  -->

<!-- ```{r} -->
<!-- # Monthly average temperatures for the years 2021 and 2022 (fake data!) -->

<!-- St_Gallen <- c(2, 2, 5, 10, 13, 21, 22, 25, 18, 14, 6, 2) -->
<!-- Lausanne  <- c(5, 6, 6, 13, 16, 23, 24, 25, 19, 16, 8, 3) -->
<!-- Basel     <- c(2, 3, 3, 10, 15, 22, 22, 24, 19, 16, 8, 3) -->

<!-- # Creating the Array -->
<!-- data_cities <- c(St_Gallen, Lausanne, Basel,  -->
<!--                  St_Gallen + 0.25*rnorm(12), Lausanne + 0.35*rnorm(12), Basel + 0.2*rnorm(12)) -->

<!-- city_array <- array(data_cities,  -->
<!--                     dim = c(12, 3, 2),  -->
<!--                     dimnames = list(months = 1:12,  -->
<!--                                     cities = c("St_Gallen", "Lausanne", "Basel"),  -->
<!--                                     year = c(2021, 2022)) -->
<!--                     ) -->
<!-- ``` -->


<!-- ## Answer the following questions: -->

<!--   1. Compute the average temperature for each city for each year  -->

<!--   2. Compute the average temperature for each city across years. -->

<!--   3. Determine which city had the highest average temperature in July 2022. -->

<!--   3. Find out the month when St. Gallen had the lowest temperature. -->

<!--   4. Compute the overall average temperature for all cities for the year. -->


<!-- ```{r} -->
<!-- # Average temperature for each city in 2021 and 2022 -->
<!-- # St. Gallen -->
<!-- for (i in 1:2) print(mean(city_array[, 1, i])) -->

<!-- # Lausanne -->
<!-- for (i in 1:2) print(mean(city_array[, 2, i])) -->

<!-- # Basel -->
<!-- for (i in 1:2) print(mean(city_array[, 3, i])) -->

<!-- # Alternatives -->
<!-- for (i in 1:3){ -->
<!--   for (j in 1:2){ -->
<!--     print(mean(city_array[, i, j])) -->
<!--   } -->
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # Compute the average temperature for each city across years. -->

<!-- # St. Gallen -->
<!-- mean(city_array[, 1, ]) -->

<!-- # Lausanne -->
<!-- mean(city_array[, 2, ]) -->

<!-- # Basel -->
<!-- mean(city_array[, 3, ]) -->


<!-- # Or (not exam relevant) -->
<!-- apply(city_array, MARGIN = 2, mean) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # Determine which city had the highest average temperature in July 2022. -->

<!-- ``` -->


<!-- print(avg_temps) -->

<!-- # b) City with highest temperature in July -->
<!-- july_temps <- city_array[7, , 1] -->
<!-- max_city <- names(which.max(july_temps)) -->
<!-- cat("City with highest avg temperature in July:", max_city, "\n") -->

<!-- # c) Month when New York had the lowest temperature -->
<!-- ny_temps <- city_array[, 1, 1] -->
<!-- min_month <- which.min(ny_temps) -->
<!-- cat("Month when New York had the lowest temperature:", min_month, "\n") -->

<!-- # d) Overall average temperature -->
<!-- overall_avg <- mean(city_array) -->
<!-- cat("Overall average temperature for all cities:", overall_avg, "\n") -->

<!-- ``` -->


