---
title: 'Data Handling: Import, Cleaning and Visualisation'
subtitle: "Exercise to lecture 4"
author: "Dr. Aurélien Sallin"
output:
  html_document:
    highlight: tango
    theme: cerulean
    mathjax: "http://mathjax.rstudio.com/latest/MathJax.js?config=TeX-MML-AM_CHTML"
  pdf_document:
    pandoc_args:
    - --filter
    - ../../code/math.py
header-includes:
- \usepackage[T1]{fontenc}
- \usepackage{hyperref}
css: ../../style/notes_hsg.css
bibliography: ../../references/datahandling.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Working with a data frame

```{r data creation, include = FALSE}
# Data Creation (to be executed by the instructor)
set.seed(123) # For reproducibility

firms <- c("FirmA", "FirmB", "FirmC", "FirmD", "FirmE")
years <- 2018:2022

# Generate sample data
data <- expand.grid(Firm = firms, Year = years)
data$Revenue <- sample(1000:5000, nrow(data), replace = TRUE)
data$Profit <- data$Revenue * runif(nrow(data), 0.1, 0.5)
data$Category <- factor(sample(c("Tech", "Health", "Finance"), nrow(data), replace = TRUE))

# Write data to .txt with ":" as separator
write.table(data, "financial_data.txt", sep = ":", row.names = FALSE)
```


## Import data
Have a look at the file `financial_data.txt` using your favorite text editor. What do you notice? Import the table using the `read.csv()` function in your environment. What does this parser do? Explore the data.frame. What is its structure? What is its dimension?

```{r warning=FALSE}
financial_data <- read.csv("financial_data.txt", sep = ":")
financial_data
```


```{r}
class(financial_data)
dim(financial_data)

str(financial_data)
names(financial_data)
dimnames(financial_data)
```

<br>

## Variable creation

Create a new variable "costs", which is the revenue - profit. [There are many ways to create a variable in a data frame, which we'll learn later in the course. Use the `$` index.]

```{r}
financial_data$costs <- financial_data$Revenue - financial_data$Profit
```

<br>

## Factor variable

Which variable is (should be) a factor? Recode this variable as a factor. What are the levels? Should we have the variable `Firm` as a factor?

```{r}
financial_data$Category <- as.factor(financial_data$Category)

levels(financial_data$Category)
```

<br>

## Nests 
Split your data using the factor variable into three data frames contained in a list. Compute the mean profit for each data frame.

```{r}
list_financial_data <- split(financial_data, financial_data$Category)

for (i in 1:length(list_financial_data)){
  print(mean(list_financial_data[[i]]$Profit))
}

# Or, using lapply
lapply(list_financial_data, function(x) mean(x$Profit))
```


<br>

Or, advanced (*and not exam relevant :-)*)
```{r message=FALSE, warning=FALSE}
# Or (advanced!) with a nested tibble and map 
library(tidyr)
library(dplyr)
library(purrr)

tibble_financial_data <- financial_data |>
  group_by(Category) |>
  nest()

map(tibble_financial_data$data, ~mean(.$Profit))

```






<!-- 2. Creating the Array: -->
<!-- Create a 3-dimensional array to store this data. The dimensions should represent: -->

<!-- Rows: Months (12 months) -->
<!-- Columns: Cities (3 cities) -->
<!-- Layers: Years (1 year for this exercise) -->
<!-- 3. Data Analysis: -->
<!-- a) Compute the average temperature for each city over the year. -->

<!-- b) Determine which city had the highest average temperature in July. -->

<!-- c) Find out the month when New York had the lowest temperature. -->

<!-- d) Compute the overall average temperature for all cities for the year. -->

<!-- ```{r} -->
<!-- # Monthly average temperatures (in Fahrenheit) for the year 2022 -->

<!-- St_Gallen <- c(2, 2, 5, 12, 15, 22, 23, 25, 18, 15, 8, 2) -->
<!-- Lausanne <- c(58, 60, 61, 65, 68, 72, 75, 76, 74, 70, 64, 59) -->
<!-- Basel <- c(27, 30, 38, 50, 60, 68, 73, 71, 64, 53, 42, 31) -->


<!-- # 2. Creating the Array -->
<!-- data_combined <- c(St_Gallen, Lausanne, Basel) -->
<!-- city_array <- array(data_combined, dim = c(12, 3, 1), dimnames = list(months = 1:12, cities = c("New York", "Los Angeles", "Basel"), year = 2022)) -->

<!-- # 3. Data Analysis -->
<!-- # a) Average temperature for each city -->
<!-- avg_temps <- apply(city_array, 2, mean) -->
<!-- print(avg_temps) -->

<!-- # b) City with highest temperature in July -->
<!-- july_temps <- city_array[7, , 1] -->
<!-- max_city <- names(which.max(july_temps)) -->
<!-- cat("City with highest avg temperature in July:", max_city, "\n") -->

<!-- # c) Month when New York had the lowest temperature -->
<!-- ny_temps <- city_array[, 1, 1] -->
<!-- min_month <- which.min(ny_temps) -->
<!-- cat("Month when New York had the lowest temperature:", min_month, "\n") -->

<!-- # d) Overall average temperature -->
<!-- overall_avg <- mean(city_array) -->
<!-- cat("Overall average temperature for all cities:", overall_avg, "\n") -->

<!-- ``` -->


