---
title: "Who is in this class?"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

<style> 


.section.sidebar {
background-color: #e6fff2;
}
selectize-control.single{
background-color: #0da861;
border-color: #0da861;

}

.navbar-inverse {
  background-color: #02934a;
  border-color: #0da861;
}
.navbar-inverse .navbar-brand {
  color: #ffffff;
}
.navbar-inverse .navbar-brand:hover,
.navbar-inverse .navbar-brand:focus {
  color: #ffffff;
  background-color: none;
}
.navbar-inverse .navbar-text {
  color: #ffffff;
}
.navbar-inverse .navbar-nav > li > a {
  color: #ffffff;
}
.navbar-inverse .navbar-nav > li > a:hover,
.navbar-inverse .navbar-nav > li > a:focus {
  color: #ffffff;
  background-color: #0da861;
}
.navbar-inverse .navbar-nav > .active > a,
.navbar-inverse .navbar-nav > .active > a:hover,
.navbar-inverse .navbar-nav > .active > a:focus {
  color: #ffffff;
  background-color: #0da861;
}
.navbar-inverse .navbar-nav > .disabled > a,
.navbar-inverse .navbar-nav > .disabled > a:hover,
.navbar-inverse .navbar-nav > .disabled > a:focus {
  color: #ffffff;
  background-color: transparent;
}
.navbar-inverse .navbar-toggle {
  border-color: transparent;
}
.navbar-inverse .navbar-toggle:hover,
.navbar-inverse .navbar-toggle:focus {
  background-color: #0da861;
}
.navbar-inverse .navbar-toggle .icon-bar {
  background-color: #ffffff;
}
.navbar-inverse .navbar-collapse,
.navbar-inverse .navbar-form {
  border-color: #1a6ecc;
}
.navbar-inverse .navbar-nav > .open > a,
.navbar-inverse .navbar-nav > .open > a:hover,
.navbar-inverse .navbar-nav > .open > a:focus {
  background-color: #0da861;
  color: #ffffff;
}
@media (max-width: 767px) {
  .navbar-inverse .navbar-nav .open .dropdown-menu > .dropdown-header {
    border-color: #0da861;
  }
  .navbar-inverse .navbar-nav .open .dropdown-menu .divider {
    background-color: #0da861;
  }
  .navbar-inverse .navbar-nav .open .dropdown-menu > li > a {
    color: #ffffff;
  }
  .navbar-inverse .navbar-nav .open .dropdown-menu > li > a:hover,
  .navbar-inverse .navbar-nav .open .dropdown-menu > li > a:focus {
    color: #ffffff;
    background-color: #0da861;
  }
  .navbar-inverse .navbar-nav .open .dropdown-menu > .active > a,
  .navbar-inverse .navbar-nav .open .dropdown-menu > .active > a:hover,
  .navbar-inverse .navbar-nav .open .dropdown-menu > .active > a:focus {
    color: #ffffff;
    background-color: #0da861;
  }
  .navbar-inverse .navbar-nav .open .dropdown-menu > .disabled > a,
  .navbar-inverse .navbar-nav .open .dropdown-menu > .disabled > a:hover,
  .navbar-inverse .navbar-nav .open .dropdown-menu > .disabled > a:focus {
    color: #ffffff;
    background-color: transparent;
  }
}
.navbar-inverse .navbar-link {
  color: #ffffff;
}
.navbar-inverse .navbar-link:hover {
  color: #ffffff;
}

.navbar-inverse .btn-link {
  color: #ffffff;
}
.navbar-inverse .btn-link:hover,
.navbar-inverse .btn-link:focus {
  color: #ffffff;
}
.navbar-inverse .btn-link[disabled]:hover,
fieldset[disabled] .navbar-inverse .btn-link:hover,
.navbar-inverse .btn-link[disabled]:focus,
fieldset[disabled] .navbar-inverse .btn-link:focus {
  color: #ffffff;
}

</style>      

```{r setup, include=FALSE}
# This code is based on Christopher Gandrud's Shiny examples:
# https://github.com/christophergandrud/datafest2017_dashboard_shiny/tree/3256125240d09bac933ed359c879036fcbeb9be5


# Load required packages. Install if needed.
library(flexdashboard)
library(gsheet)
library(ggmap)
library(dplyr)
library(leaflet)
library(sp)
library(rworldmap)
library(networkD3)
library(geosphere)
library(plotly)
library(shiny)


# fix vars
#  API key needed here
# KEY <- readLines("../../admin/2021/keys/datahandling_survey_static.txt") # key UM
KEY <- "AIzaSyASzCqktyaJwksvy9Oi3i50gHKuiAFOP5g"

# Download survey data
# Download data
URL <- 'https://docs.google.com/spreadsheets/d/13jZFfQHdqN5fI4PqGZZCvHGSgPbDPKBhdSmQ4WClyME/edit?gid=0#gid=0'
student_data <- gsheet2tbl(URL)

# Simplify variable names
names(student_data) <- c('home_town', 'used_R', 'literacy', "major", "id")

# Relevel factors
r_skills <- 1:10
student_data$literacy <- factor(student_data$literacy, levels = r_skills)

# Majors
majors <- c('VWL', 'BWL', 'BIA', 'BLaw', 'BLE','other')
student_data$major <- factor(student_data$major, levels = majors)

# Used R
student_data$used_R <- factor(student_data$used_R, levels = c('TRUE', 'FALSE'))

# Find home_town longitude and latitude -----
home_towns <- student_data$home_town
home_towns <- home_towns[!is.na(home_towns)]
register_google(key = KEY)
home_town_coords <- geocode(home_towns) # problem: might reach limit!

home_town_coords <- na.omit(home_town_coords)
```


Geography
===========


Column {data-width=650}
-----------------------------------------------------------------------

### Map of home_towns

```{r home_town-map}
# Map student home_towns
leaflet() %>% 
  addTiles() %>%
  addCircleMarkers(data = home_town_coords, lng = ~lon, lat = ~lat, color = "#02934a")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Close (<= 100 km) home_towns Network

```{r}
# Maximum distance for relationship
max_distance <- 100000 # in meters

# Find distances between each home town
distances <- vector()
for (i in 1:nrow(home_town_coords)) {
    temp_indv <- home_town_coords[i, ]
    
    temp_distances <- distm(home_town_coords, temp_indv) %>% data.frame
    temp_distances$id <- i
    temp_distances$other_id <- 1:nrow(home_town_coords)
    names(temp_distances)[1] <- 'distance'
    distances <- rbind(distances, temp_distances)
}
distances <- distances[, c(2, 3, 1)]

# Remove self-self edges
distances <- subset(distances, id != other_id)

# Keep only nodes that are within 1,000 km of each other
if (any(distances$distance <= max_distance)) {
  distances <- subset(distances, distance <= max_distance)
}

# Plot network
simpleNetwork(distances, zoom = TRUE, linkColour = "#02934a", nodeColour = "#02934a", charge = 0)
```

### Home Countries

```{r}
# Function to find countries that each home_town is in
coords2country = function(points)
{  
    # Function from: http://stackoverflow.com/a/14342127
    countriesSP <- getMap(resolution='low')

    #setting CRS directly to that from rworldmap
    pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  

    # use 'over' to get indices of the Polygons object containing each point 
    indices = over(pointsSP, countriesSP)

    # return the ADMIN names of each country
    indices$ISO3  
}
countries <- coords2country(na.omit(home_town_coords))

# Find country counts
countries_count <- table(countries)
countries_count <- countries_count[countries_count > 0] 
countries_count <- data.frame(countries_count)

# Plot counts
plot_ly(data = countries_count, x = ~countries, y = ~Freq, color = I("#02934a")) %>%
    layout(xaxis = list(title = ''),
           yaxis = list(title = 'Frequency'))
```



Computer Literacy
=================



Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("majors", label = "Select Participants' Major",
            choices = majors, selected = '')
```

Column
------------------------------------------

```{r}
renderPlot({
    # Subset data
    data_sub <- subset(student_data, major == input$majors)
    
    # Find major counts
    r_exp <- table(data_sub$literacy)
    
    # Bar plot
    barplot(r_exp)
    
})
```


Who has already used R?
=======================

Column {data-width=150}
-----------------------------------------------------------------------

```{r, fig.width=10, fig.height=7}
renderPlot({
    # Find major counts
    r_exp <- table(student_data$used_R)

    # Calculate percentages
    r_exp_percents <- prop.table(r_exp) * 100
    r_exp_percents <- as.data.frame(r_exp_percents)
    names(r_exp_percents) <- c("Used R", "Pct")

    # Bar plot
    ggplot(r_exp_percents, aes(x = `Used R`, y = Pct)) +
        geom_bar(stat = "identity", fill = "#02934a") +
        geom_text(aes(y = Pct/2, label = paste0(round(Pct, 1), "%")), 
                  size = 3, color = "white") +
        theme_classic()  
    
})
```

Column {data-width=230}
-----------------------------------------------------------------------


Analytics
=================



Column 
-----------------------------------------------------------------------

```{r}
renderPlot({
    
     # Find major counts
     df <- student_data %>% 
       select(major, literacy) %>% 
          group_by(major) %>% 
          summarise(mean_ability = mean(as.numeric(literacy), na.rm = T))
    
    # Bar plot
     student_data %>% 
       select(major, literacy, home_town) %>% 
          na.omit() %>% 
          left_join(df) %>% 
          ggplot(aes(x = major, y = literacy),) +
          geom_point(stat = "identity", color = "#02934a") +
          geom_point(aes(y = mean_ability), color = "red", shape=1,size=3) +
          theme_classic()
    
})
```







